---
title: "IGN Video Game Reviews"
author: "Ellen A Savoye"
date: "July 12, 2018"
output: pdf_document
subtitle: Capstone
---
***
  
  
# Introduction
Over the last 20 years, a plethora of video games have been released in an ever-growing market of gaming consoles. In tandem with the creation of these games, a website called IGN began publishing reviews and ratings of the games mentioned above. These reviews and resulting ratings can be analyzed to create insights that would be useful to console and video game creators. With analysis, console and video game creators would be able to determine their market/popularity standing in comparison to other consoles and games. Furthermore, they would be able to determine if there is a particular genre that needs more development to increase their rating and market standpoint. Sourced from Kaggle (<https://www.kaggle.com/egrinstein/20-years-of-games>) and IGN (<http://ign.com/games/reviews>) via a crawl, the data consists of 20 years worth of video game data.

# Caveats
The data does not contain any financial information relating to the volume of games sold or the monetary amount it sold for.  Insights are curated based on the number of games released, their rating, the genre, and the console the game was released on. Some games, like 'Gears of War,' have been released on multiple consoles. I have not adjusted the data to constrain these types of games down to one platform.

# The Data
The data, 20 years' worth of IGN game reviews, consists of 18,625 records. The raw dataset had ten columns (listed below). 

|Variable|Description|
|----|----|
|score_phrase| Phrase given to describe the overall score
|title| Game title 
|url| IGN Game URL
|platform| Game Console
|genre| Video game genre
|score| Overall rating for the video game
|editors_choice| Editor Recommended (Y/N) 
|release_year| Year of game release
|release_month| Month of game release
|release_day| Day of game release

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r game_reviews data, echo=FALSE, include=FALSE}

library("knitr")
library("readr")
library("dplyr")
library("tidyr")
library("RColorBrewer")
library("ggplot2")
library("caTools")
library("randomForest")
library("rpart")
library("rpart.plot")
library(caret)

IGN_data <- read_csv("ign_original.csv")
```

```{r game_reviews table, echo=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80), eval=FALSE}
# preview of the data set
pander::pander(IGN_data[1:5, ], caption="IGN Data")

```

## Data Wrangling

To take a proper look at the data, I loaded the original dataset as a CSV file and the necessary libraries. Of the variables available for use, score_phrase, platform, score, genre, editors_choice, release_year, release_month, and release_day are the ones I used in my analysis. As such, I analyzed them for missing values, outliers, and whether or not the number of distinct factors in each was usable. Editors_choice, score_phrase, and score did not need cleaning. However, when checking release_year, I noticed an outlier titled "The Walking Dead: The Game -- Episode 1: A New Day". This record had a release date of 1/1/1970. Given the dataset is spanning 1996 - 2016, I chose to correct the outlier to the correct release date of 4/24/2012.

```{r show_table, tidy=TRUE, max.print = 10, tidy.opts=list(width.cutoff=60)}
head(tbl_df(IGN_data), 5)
```

```{r check_release_year, tidy=TRUE, tidy.opts=list(width.cutoff=60), max.print = 10}
# Release year is supposed to be higher than 1995

head(IGN_data %>% distinct(release_year),5) %>% 
arrange(release_year)

IGN_data[IGN_data$release_year == "1970", ]

IGN_data <- IGN_data %>% mutate(release_year = if_else(title == "The Walking Dead: The Game -- Episode 1: A New Day", as.integer(2012), release_year)) %>% 
mutate(release_month = if_else(title == "The Walking Dead: The Game -- Episode 1: A New Day", as.integer(4), release_month)) %>% 
mutate(release_day = if_else(title == "The Walking Dead: The Game -- Episode 1: A New Day", as.integer(24), release_day))

```

With the outlier corrected, platform and genre variables remained. The original platform variable consisted of 59 distinct factors. Because platform spanned multiple generations of systems (e.g., PlayStation 1-3) and because not all manufacturers kept system naming consistent, I chose to combine the values into a condensed version based on system name/manufacturer and created a new variable named platform_group. To do so, I loaded a 'platform map' CSV file to merge the new platform_group variable onto the original dataset. After comparing the original platform variable against the new platform_group to ensure no misplaced systems, I moved onto the genre variable.

```{r condense_platform_list, results='hide', tidy=TRUE}
# 59 variables in original platform column
IGN_data %>% distinct(platform) %>% 
arrange(platform)

Platform_Map <- read_csv("platform_map.csv") 

IGN_data_v2 <- IGN_data %>% mutate_at(vars(platform),funs(as.factor(as.character(.))))

IGN_data_v2 <- IGN_data_v2 %>% left_join(Platform_Map, by = c("platform" = "platform"))
```

```{r platform_group_list, tidy=TRUE, max.print = 10}
IGN_data_v2 %>% group_by(platform, platform_group) %>% 
summarise( n_distinct(platform_group))
```

Similar to the platform variable, the genre variable has a multitude of factors which makes intelligent analysis a bit difficult. There are 113 unique genres within the field. I chose my grouping based on an overall description (e.g., Sports, Cards, Action, etc.) given the numerous distinct factors. Before cleaning up the column, I checked for any blank cells. Out of 18,625 observations, 36 do not have a genre which is .19%. Due to the unpopulated records being less than 1% of the overall genre column, I chose not to populate them but instead mapped them to 'Other.' To map genre, I loaded a 'genre map' CSV file to merge the new genre_group variable onto the original dataset. In doing so, I brought the number of unique genres from 113 to 21.

```{r check_genre, tidy=TRUE, results='hide'}
# Check for blanks in genre column

IGN_data_v2 %>% distinct(genre) %>% 
  arrange(genre)
```

```{r check_genre_blanks, tidy=TRUE, max.print = 10}
group_by(IGN_data_v2[is.na(IGN_data_v2$genre), ])
```

```{r group_genre_list, results='hide', tidy=TRUE}
# 113 unique factors in genre column 

IGN_data_v2 %>% distinct(genre) %>% 
arrange(genre)

Genre_Map <- read.csv("genre_map.csv")

Genre_Map <- Genre_Map %>% mutate_at(vars(genre),funs(as.character(as.factor(.))))

IGN_data_v2 <- IGN_data_v2 %>% left_join(Genre_Map, by = c("genre" = "genre"))
```
```{r genre_group_list, tidy=TRUE, max.print = 10}
unique(IGN_data_v2$genre_group)
length(unique(IGN_data_v2$genre_group))
```

After cleaning up the variables that I will be using in my analysis, I wrote the wrangled data to a new file called "ign_clean.csv" for further use in creating insights.

```{r cleaned_code_CSV, results='hide', tidy=TRUE}

write_csv(IGN_data_v2, "ign_clean.csv")
IGN_data_cleaned <- read.csv("ign_clean.csv")

IGN_data_cleaned <- IGN_data_cleaned[,-c(1,4)]

```

***
# Exploratory Data Analysis

## Genre

Using the condensed genre field, I'm looking to see what the top genres are regarding the number of video game releases and whether or not the scores correspond to the top genres.

#### Top Genres
```{r top_genres, tidy=TRUE}

top_genres <- IGN_data_cleaned %>%
  group_by(genre_group) %>%
  dplyr::summarize(genres_count = n()) %>%
  arrange(desc(genres_count))

top_genres$genre_group <- factor(top_genres$genre_group, levels = top_genres$genre_group[order(top_genres$genres_count)])

colourCount = length(unique(top_genres$genre_group))
fill_purple <- colorRampPalette(brewer.pal(9,"BuPu"))

top_genres %>% 
  filter(genre_group != "NA") %>% 
  ggplot(aes(x = genre_group, y = genres_count, fill = genre_group)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  geom_text(aes(label = genres_count), size = 2.5, color = "black", hjust = -.5) +
  labs(x = "Genre", y = "Number of Games", title = "Genres of Video Games for 1996-2016") +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
  ylim(0, max(top_genres$genres_count + 100)) +
  scale_fill_manual(values = fill_purple(colourCount))

```

Cumulatively, the top genre is Action followed by Sports, Shooter, Adventure, and Racing. The number of Action games is more than double the next genre, Sports. As I mentioned in the caveats section, this may be due to some games applying across platforms and therefore counted multiple times. Now that we have the top genres, we can see if there are any insights to glean.

#### Top Genres by Score

Due to a little more than half of the genres having a count of 500+, I'm using 500 as my minimum number of games to filter my data. However, given that 'other' is a catch-all bucket, the minimum will instead be 550 to exclude 'other.'


```{r top_genres_scores, tidy=TRUE}

top_genre_scores <- IGN_data_cleaned %>%
group_by(genre_group) %>%
dplyr::summarize(sum_genres_count = n()) %>%
arrange(desc(sum_genres_count)) %>% 
filter(sum_genres_count > 550) 
top_genre_scores <- top_genre_scores[,1]

genre_scores_df <- IGN_data_cleaned[IGN_data_cleaned$genre_group %in% top_genre_scores$genre_group,]

colourCount = length(unique(top_genres$genre_group))
fill_purple <- colorRampPalette(brewer.pal(9,"BuPu"))

genre_scores_df %>% ggplot(aes(x = genre_group, y = score, fill = genre_group)) +
  geom_boxplot(alpha = 1) +
  labs(x = "Top Genres", y = "Score", title = "Distribution of Scores by Top Genres for 1996-2016") +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = -45, hjust = 0)) +
  geom_hline(aes(yintercept = median(genre_scores_df$score)), color = "red") +
  scale_fill_manual(values = fill_purple(colourCount))

subset(genre_scores_df, genre_scores_df$score < 1)

```

When looking at the distribution of scores for the top 11 genres, we can see that RPG has higher ratings than the other genres even though it sits in sixth place for volume of games released. Given the dearth of games released in the Action genre, I expected it to have the higher scores across the board from a slight volume influenced stand-point. Action, Adventure, and Racing have lower scores with Racing having the largest IQR. Both Action and Racing have games with a rating less than 1. Two of the games are in Racing under while the third is in Action.


## Platforms

I want to see which platform reigned supreme over the last twenty years and if that holds true when looking at the past five years.

#### Top Platforms
```{r top_platforms, tidy=TRUE}

top_platform <- IGN_data_cleaned %>%
group_by(platform_group) %>%
dplyr::summarize(platform_count = n()) %>%
arrange(desc(platform_count))

top_platform$platform_group <- factor(top_platform$platform_group, levels = top_platform$platform_group[order(top_platform$platform_count)])

colourCount = length(unique(top_platform$platform_group))
fill_purple <- colorRampPalette(brewer.pal(9,"PuRd"))

top_platform %>% 
  filter(platform_group != "NA") %>% 
  ggplot(aes(x = platform_group, y = platform_count, fill = platform_group)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  geom_text(aes(label = platform_count), size = 2.5, color = "black", hjust = -.5) +
  labs(x = "Platform", y = "Number of Platforms", title = "Video Game Platforms for 1996-2016") +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
  ylim(0, max(top_platform$platform_count + 100)) +
  scale_fill_manual(values = fill_purple(colourCount))

```

While some of these platforms have had multiple versions/evolutions over the last 20 years, Nintendo for example, they have been grouped for a more straightforward analysis on the major gaming platforms. Even though PlayStation launched after Nintendo, PlayStation is still the top system followed by Nintendo, Windows, and Xbox. Given the previous graph consisted of all 20 years worth of data, I put together the last five years of data into a chart to see if the top systems remained consistent.

#### Top Platforms for Past 5 Years

```{r top_platforms_score, tidy=TRUE}

platform_years <- IGN_data_cleaned %>%
  filter(release_year >= 2012) %>%
  group_by(release_year,platform_group) %>%
  dplyr::summarize(sum_platform_count = n()) %>%
  arrange(desc(release_year),desc(sum_platform_count)) %>% 
  top_n(5, sum_platform_count)
  
platform_years %>% ggplot(aes(x = platform_group, y = sum_platform_count, fill = release_year)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ release_year, nrow = 5) + 
  coord_flip() +
  geom_segment(aes(x=platform_group,xend=platform_group, y=0, yend=sum_platform_count)) +
  labs(title = "Top Platforms (2012-2016)", x='',y='Count') +
  geom_text(aes(label=sum_platform_count), hjust = -1, size = 3) +
  theme(legend.position = 'none', plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~ release_year, nrow = 5, scales = 'free', strip.position = 'right') +
  ylim(0, max(platform_years$sum_platform_count + 10)) 

```

For 2012-2014, Playstation held down the top spot which is consistent with all years combined. However, 2015-2016, Windows is the top platform. So even though Playstation holds the top place in the number of games released overall, it doesn't necessarily hold true on a year-by-year basis.

#### Top Platforms by Score

Taking the top 5 platforms found in our "Top Platforms" exhibit above, I want to see how the scores correspond to platform volume.


```{r top_platforms_scores, tidy=TRUE}

top_platform_scores <- IGN_data_cleaned %>%
  group_by(platform_group) %>%
  dplyr::summarize(sum_platforms_count = n()) %>%
  arrange(desc(sum_platforms_count)) %>% 
  top_n(5, sum_platforms_count)
top_platform_scores <- top_platform_scores[,1]

platform_scores_df <- IGN_data_cleaned[IGN_data_cleaned$platform_group %in% top_platform_scores$platform_group,]

colourCount = length(unique(top_platform_scores$platform_group))
fill_purple <- colorRampPalette(brewer.pal(9,"PuRd"))

platform_scores_df %>% ggplot(aes(x = platform_group, y = score, fill = platform_group)) +
  geom_boxplot(alpha = 1) +
  labs(x = "Top platforms", y = "Score", title = "Distribution of Scores by Top Platforms for 1996-2016") +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = -45, hjust = 0)) +
  geom_hline(aes(yintercept = median(genre_scores_df$score)), color = "green") +
  scale_fill_manual(values = fill_purple(colourCount))

```

Given the number of game releases that Playstation has had, there was some expectation for a better score performance. However, this doesn't appear to be the case. Both Xbox and Windows have higher scores for the 25th, 50th, and 75th IQR values. Windows has the top number of released games for 2015 and 2016, so the higher scores are not wholly a surprise. One thing I've noticed personally is a shift towards computer-based, Windows, games due to the ability to fine-tune graphics cards and speed for better gaming performance.

### Important Dates
We want to see if a particular month and year stand out as significant. To do so, I looked at the month and year together in a grid.

```{r Date_Year, tidy=TRUE}

Mon_Yr_Ct <- IGN_data_cleaned %>%
  filter(release_year > 1996) %>% 
  group_by(release_month, release_year) %>%
  dplyr::summarize(games_per_mon = n()) %>%
  arrange(desc(games_per_mon))
Mon_Yr_Ct %>% 
  mutate(release_month_2 = as.Date(paste0(release_month,"-01"),"%m-%d")) %>% 
  ggplot(aes( x = release_month_2, y = games_per_mon, fill = release_month_2)) +
  geom_bar(stat = "identity") +
  facet_wrap( ~release_year, ncol = 5) + 
  labs(title = "Game by Release Month and Year", y = "Game Release per Month", x = "Month") +
  theme(legend.position = 'none', plot.title = element_text(hjust = 0.5), axis.text.x = element_text(angle = -45, hjust = 0)) +
  scale_x_date(date_labels = "%b")
  
```

From 1997 - 2008, the number of games released continuously increases, with 2008 having the most reviewed and published in video game history. In older years, there is a more significant amount of released video games in October and November over other months. The more recent years have more consistent releases over the course of the year. The decreased number of games released could be due to video games becoming significantly more intricate and graphics intensive. With the capabilities of T.V. and graphics cards now, immersive video game graphics seem like a must-have for a game to be released.

***

# Predictive Models

#### Subsetting the Data

Before any ventures into modeling, a data set was created to remove any unnecessary features and any blank records. 36 records with a 'blank' genre were removed. Given that some of the kept features are categorical, conversion to binary/dummy variables was needed using a model matrix. I chose not to combine release day, month, and year into a combined date field as I don't believe it would have given any valuable insight as unique a combined value. 


```{r subset_data, tidy=TRUE}
# Check for blank/NA records
sapply(IGN_data_cleaned, function(x) sum(is.na(x)))
IGN_DC_PredM_WONA <- na.omit(IGN_data_cleaned)
IGN_DC_PredM <- IGN_DC_PredM_WONA[, c(-1, -2, -3, -5)]

#Let convert time related variable to factor so we can binarize them
IGN_DC_PredM$release_day   <- as.factor(IGN_DC_PredM$release_day)
IGN_DC_PredM$release_month <- as.factor(IGN_DC_PredM$release_month)
IGN_DC_PredM$release_year  <- as.factor(IGN_DC_PredM$release_year)

# str(IGN_DC_PredM)

IGN_MM <- model.matrix(~ editors_choice + platform_group + genre_group + release_day + release_month + release_day, data=IGN_DC_PredM)

# dim(IGN_MM)

#drop near zero predictor
nzr <- nearZeroVar(IGN_MM)
# nzr
IGN_MM <- IGN_MM[, -nzr]

# dim(IGN_MM)

#drop highly correlated features
corel <- cor(IGN_MM)
highCorrel <- findCorrelation(corel, cutoff = 0.85) #no correlation above 85%. Great!
# highCorrel

#Now let's combine our data for regression analysis.
reg_data <- cbind(IGN_MM, score = IGN_DC_PredM$score)

# Hmisc::describe(reg_data)
```


#### Linear Regression

```{r sig_features, tidy=TRUE}
#convert our matrix data to a data frame for ease of use
reg_data <- as.data.frame(reg_data)

#make sure column names are compatible with R naming convention
colnames(reg_data) <- make.names(colnames(reg_data))

ScoreReg <- lm(score ~., data = reg_data)
summary(ScoreReg)

reg_data$platform_groupApple <- NULL
reg_data$genre_groupAdventure <- NULL
reg_data$genre_groupSports <- NULL
reg_data$release_month5 <- NULL
reg_data$release_month12 <- NULL

ScoreReg <- lm(score ~., data = reg_data)
summary(ScoreReg)

reg_data$genre_groupShooter <- NULL
reg_data$release_month7 <- NULL

ScoreReg <- lm(score ~., data = reg_data)
summary(ScoreReg)
```

The first iteration of the linear regression showed multicollinearity between score and score phrase with a multiple R-squared of 0.9726. After removing score_phrase, R-squared drops to 0.3187. In an attempt to improve R-squared, near zero variance predictors (NZP) were identified and discarded because they are non-informative and tend to occur when breaking categorical variables into dummy variables, as was done above. After removing NZP and checking for high collinearity, a few iterations of the linear model were run to narrow down which features are not significant. A few of the fields removed are platform_groupApple, genre_groupAdventure, and genre_groupSports. After eliminating non-significant fields, the adjusted R-squared is now 0.3127. 

#### CART/Random Forest

```{r random_forest, tidy=TRUE}

set.seed(1000)

spl <- sample.split(reg_data$score, SplitRatio = 0.8)
Train <- subset(reg_data, spl == TRUE)
Test <- subset(reg_data, spl == FALSE)

# CART
IGN_Tree <- rpart(score ~., data = Train, method = "class", minbucket = 25)
prp(IGN_Tree)

IGN_Tree_v2 <- rpart(score ~., data = Train, method = "class", control = rpart.control(minbucket = 25, cp = 0.0025))
prp(IGN_Tree_v2)

# Random Forest
IGN_Forest <- randomForest(score ~., data = Train, ntree = 500)

IGN_Forest

preds_f <- predict(IGN_Forest, Test)
#R-square of the final RF model -> 32.69 %
cor(preds_f, Test$score)^2

preds_t <- predict(IGN_Tree, Test)
#R-square of the final 1 node Tree model -> ~30.4 %
cor(preds_t, Test$score)^2

preds_t_v2 <- predict(IGN_Tree_v2, Test)
#R-square of the final multi-node Tree model -> ~30.4 %
cor(preds_t_v2, Test$score)^2

```

When looking at the first tree, the only node shown is for editor's choice. If editor's choice is higher than 0.5, it is no or 1. One important note is that a high score can imply an editor's choice designation. However, a high score, say 8 or higher, does not automatically receive an editor's choice designation. Using the complexity parameter (cp), we can force more nodes to appear with cp = .0025. Looking at the tree, we see that editor's choice has remained as the first node. The first tree produces an R-squared of 30.4% which is similar to the linear regression model. When forcing the tree to have multiple nodes, the same R-squared calculation doesn't work as nicely nor is it the most efficient.

The random forest model, built on the training dataset, appears to be the best model based on a comparison of R-squared at 32.69% versus 30.4% (CART), and 31.27% (linear regression).

***

# Conclusion

Between our exploratory data analysis and machine learning models, there are a few takeaways. While the three models don't have high R-squared, this doesn't mean our models are inadequate. The predictor values in each are significant which we can still make important conclusions on. Based on the predictor values, we can see that while specific platforms are significant in their effect on the score, the relationship is negative. Genres RPG and Strategy along with the significant release months have a positive relationship with our score variable. One thing to keep in mind is the human factor. Human opinion influenced the ratings for each game in our source dataset which can make it more challenging to create a precise model. 

The human aspect is one limitation to this analysis. Another limitation is the lack of detail in what went into the determination of scores, other than genre and console, for each review: i.e., graphics rating, storyline, etc. 

Based on my findings, I would recommend an additional study be performed incorporating more information regarding score determination as well as financial information relating to each game: sales volume, cost of each game, etc. The additional information will, theoretically, provide more detail to aid in accurately predicting video game rating and how they stack up to each other in the market.
