---
output:
  html_document: default
---

# Capstone Proposal 
### 20 Years of IGN Game Reviews
##### By Ellen Savoye

***

#### **The Problem**
*What is the problem I want to solve?*

Over the last 20 years, a plethora of video games have been released in an ever growing market of gaming consoles. Given the diversity, I would like to determine if, by year and over the years, there is a particular console that is the most popular by use and by average game rating. In addition, I would like to determine if, by year and over the years within reason, there is a particular video game genre that is heralded as the favorite by rating and amount of games within that genre with a positive rating.

***

#### **The Client**
*Who is your client and why do they care about this problem? In other words, what will your client do or decide based on your analysis that they wouldn’t have otherwise?*

The client of this analysis would be console and video game creators. With the results, the client would be able to determine their market/popularity standing in comparison to other consoles. Furthermore, the client would be able to determine if there is a particular genre that needs more development in order to potentially increase their rating and market standpoint. 

***

#### **The Data**
*What data are you going to use for this? How will you acquire this data?*

The dataset for this project, via a crawl for retrieval, comes from <http://ign.com/games/reviews>. The data consists of 20 years' worth of IGN game reviews with 18,625 records. 

|Variable|Description|
|----|----|
|score_phrase| Phrase given to describe overall score
|title| Game title 
|url| IGN Game URL
|platform| Game Console
|genre| Video game genre
|score| Overall rating for video game
|editors_choice| Editor Recommended (Y/N) 
|release_year| Year of game release
|release_month| Month of game release
|release_day| Day of game release


```{r game_reviews data, echo=FALSE, include=FALSE}
#install.packages("kableExtra")

library(knitr)
library(readr)
library(dplyr)

game_reviews <- read_csv("ign.csv")

```

```{r game_reviews table, echo=FALSE, results='asis'}
# preview of the data set

knitr::kable(game_reviews[1:5, ], caption="IGN Data")

```

***

#### **The Approach**
*In brief, outline your approach to solving this problem (knowing that this might
change later)*

My approach can be broken out into the following sections:

1. Data wrangling and cleaning
    + Take into account any missing values or outliers
    + Are there any discontinued consoles? How should they be accounted for?
2. Exploratory Data Analysis
    + Use a combination of inferential statistics and data visualization to identify trends between game consoles and video game ratings
        + Determine potentially significant variables
        + Identify trends and correlations between variables
3. Machine Learning
    + Dig deeper into the correlation between consoles and game ratings using classification techniques and/or regression models
4. Data visualization and report out
    + Compile all relevant information into deliverables (listed below)
  
***

#### **Deliverables**
*What are your deliverables?*

My deliverables will consist of a report on my findings, a slide deck, and the corresponding R code used in analyzing the data. In addition, the aforementioned deliverables will be submitted and published on GitHub.


