---
title: 'IGN Data: Data Wrangling'
author: 'by Ellen A. Savoye'
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---
***
<br>

The IGN dataset sourced from Kaggle (<https://www.kaggle.com/egrinstein/20-years-of-games>) and IGN (<http://ign.com/games/reviews>), via a crawl, consists of 20 years worth of video-game data. To take a proper look at the data, I loaded the original dataset as a CSV file and the necessary libraries.

```{r load_data, echo=FALSE, include=FALSE}
# Load libraries and read IGN CSV file

library("dplyr")
library("tidyr")
library("knitr")

IGN_data <- read.csv("ign_original.csv")
```

```{r show_table, results='hide'}
tbl_df(IGN_data)
```

<br>

Of the variables available for use, score_phrase, platform, score, genre, editors_choice, release_year, release_month, and release_day are the ones I am using in my analysis. As such, I analyzed them for missing values, outliers, and whether or not the number of distinct factors in each was usable. Editors_choice, score_phrase, and score did not need cleaning. However, when checking release_year, I noticed an outlier titled "The Walking Dead: The Game -- Episode 1: A New Day". This record had a release date of 1/1/1970. Given the dataset is spanning 1996 - 2016, I chose to correct the outlier to the correct release date of 4/24/2012.

```{r check_release_year, results='hide', tidy=TRUE, tidy.opts=list(width.cutoff=60)}
# Release year is supposed to be higher than 1995

IGN_data %>% distinct(release_year) %>% 
  arrange(release_year)

IGN_data[IGN_data$release_year == "1970", ]

IGN_data <- IGN_data %>% mutate(release_year = if_else(title == "The Walking Dead: The Game -- Episode 1: A New Day", as.integer(2012), release_year)) %>% 
  mutate(release_month = if_else(title == "The Walking Dead: The Game -- Episode 1: A New Day", as.integer(4), release_month)) %>% 
  mutate(release_day = if_else(title == "The Walking Dead: The Game -- Episode 1: A New Day", as.integer(24), release_day))

```

<br>

With the outlier corrected, platform and genre variables remained. The original platform variable consisted of 59 distinct factors. Because platform spanned multiple generations of systems (e.g., PlayStation 1-3) and because not all manufacturers kept system naming consistent, I chose to combine the values into a condensed version based on system name/manufacturer and created a new variable named platform_group. To do so, I loaded a 'platform map' CSV file to merge the new platform_group variable onto the original dataset. After comparing the original platform variable against the new platform_group to ensure no misplaced systems, I moved onto the genre variable.

```{r condense_platform_list, results='hide', tidy=TRUE}
# 59 variables in original platform column

IGN_data %>% distinct(platform) %>% 
  arrange(platform)

Platform_Map <- read.csv("platform_map.csv")

IGN_data <- IGN_data %>% left_join(Platform_Map, by = c("platform" = "platform"))

IGN_data %>% group_by(platform, platform_group) %>% 
  summarise(n_distinct(platform))

```

<br>

Similar to the platform variable, the genre variable has a multitude of factors which makes intelligent analysis a bit difficult. There are 113 unique genres within the field. I chose my grouping based on an overall description (e.g., Sports, Cards, Action, etc.) given the numerous distinct factors. Before cleaning up the column, I checked for any blank cells. Out of 18,625 observations, 36 do not have a genre which is .19%. Due to the blank records being less than 1% of the overall genre column, I chose not to populate them but instead mapped them to 'Other.' To map genre, I loaded a 'genre map' CSV file to merge the new genre_group variable onto the original dataset. In doing so, I brought the number of unique genres from 113 to 21.

```{r check_genre, results='hide', tidy=TRUE}
# Check for blanks in genre column

IGN_data %>% distinct(genre) %>% 
  arrange(genre)

IGN_data[IGN_data$genre == "", ]
```

```{r group_genre_list, results='hide', tidy=TRUE}

# 113 unique factors in genre column 

IGN_data %>% distinct(genre) %>% 
  arrange(genre)

Genre_Map <- read.csv("genre_map.csv")

IGN_data <- IGN_data %>% left_join(Genre_Map, by = c("genre" = "genre"))
                                                  
unique(IGN_data$genre_group)
```

<br>

After cleaning up the variables that I will be using in my analysis, I wrote the wrangled data to a new file called "ign_clean.csv" for further use later in the course.

```{r cleaned_code_CSV, results='hide', tidy=TRUE}

write.csv(IGN_data, "ign_clean.csv")

```

















