---
title: "IGN Data: Data Wrangling"
output: html_notebook
---

##### by Ellen A. Savoye

The IGN data set is sourced from Kaggle (<https://www.kaggle.com/egrinstein/20-years-of-games>) and IGN (<http://ign.com/games/reviews>) via a crawl. In order to take a proper look at the data, I loaded the original dataset as a CSV file and the necessary libraries.

```{r load_data}
# Load libraries and read IGN CSV file

library("dplyr")
library("tidyr")
library("knitr")

IGN_data <- read.csv("ign_original.csv")

tbl_df(IGN_data)

```

Of the variables available for use, score_phrase, platform, score, genre, editors_choice, release_year, release_month, and release_day are the ones I will be using in my analysis. As such, they needed to be checked for missing values, outliers, and whether or not the amount of distinct factors in each was usable. Editors_choice, score_phrase, and score did not need to be cleaned. However when checking release_year, I noticed an outlier titled "The Walking Dead: The Game -- Episode 1: A New Day". This record had a release date of 1/1/1970. Given the dataset is spanning 1996 - 2016, I chose to correct the outlier to the correct release date of 4/24/2012.

```{r check_release_year, results='hold'}
# Release year is supposed to be greater than 1995

IGN_data %>% distinct(release_year) %>% 
  arrange(release_year)

IGN_data[IGN_data$release_year == "1970", ]

IGN_data <- IGN_data %>% mutate(release_year = if_else(title == "The Walking Dead: The Game -- Episode 1: A New Day", as.integer(2012), release_year)) %>% 
  mutate(release_month = if_else(title == "The Walking Dead: The Game -- Episode 1: A New Day", as.integer(4), release_month)) %>% 
  mutate(release_day = if_else(title == "The Walking Dead: The Game -- Episode 1: A New Day", as.integer(24), release_day))

```

With the outlier corrected, platform and genre remained to be cleaned. The original platform variable consisted of 59 distinct factors. Because platform spanned multiple generations of systems (e.g. Playstation 1-3) and because not all manufacturers kept system naming consistent, I chose to combine the values into a condensed version based largely on system name/manufacturer and created a new variable named platform_group. After comparing the original platform variable against the new platform_group to ensure no misplaced systems, I moved onto genre.

```{r condense_platform_list}
# 59 variables in original platform column

IGN_data %>% distinct(platform) %>% 
  arrange(platform)

IGN_data$platform_group <- ifelse(grepl("Playstation", IGN_data$platform, ignore.case = TRUE), "Playstation", 
                                  ifelse(grepl("Game Boy", IGN_data$platform, ignore.case = TRUE), "Game Boy",
                                          ifelse(grepl("Sega", IGN_data$platform, ignore.case = TRUE) | grepl("Genesis", IGN_data$platform, ignore.case = TRUE) | grepl("Dreamcast", IGN_data$platform, ignore.case = TRUE) | grepl("Master System", IGN_data$platform, ignore.case = TRUE) | grepl("Saturn", IGN_data$platform, ignore.case = TRUE), "Sega",
                                                  ifelse(grepl("Nintendo", IGN_data$platform, ignore.case = TRUE) | grepl("NES", IGN_data$platform, ignore.case = TRUE) | grepl("Wii", IGN_data$platform, ignore.case = TRUE) | grepl("GameCube", IGN_data$platform, ignore.case = TRUE), "Nintendo",
                                                          ifelse(grepl("Xbox", IGN_data$platform, ignore.case = TRUE), "Xbox",
                                                                  ifelse(grepl("Iphone", IGN_data$platform, ignore.case = TRUE) | grepl("Ipod", IGN_data$platform, ignore.case = TRUE) | grepl("Ipad", IGN_data$platform, ignore.case = TRUE) | grepl("Macintosh", IGN_data$platform, ignore.case = TRUE), "Apple",
                                                                          ifelse(grepl("Atari", IGN_data$platform, ignore.case = TRUE) | grepl("Lynx", IGN_data$platform, ignore.case = TRUE), "Atari",
                                                                                  ifelse(grepl("Windows", IGN_data$platform, ignore.case = TRUE) | grepl("PC", IGN_data$platform, ignore.case = TRUE), "Windows",
                                                                                          ifelse(grepl("NeoGeo", IGN_data$platform, ignore.case = TRUE), "NeoGeo",
                                                                                                  ifelse(grepl("Ouya", IGN_data$platform, ignore.case = TRUE) | grepl("Android", IGN_data$platform, ignore.case = TRUE), "Android",
                                                                                                          ifelse(grepl("SteamOS", IGN_data$platform, ignore.case = TRUE) | grepl("Linux", IGN_data$platform, ignore.case = TRUE), "Linux",
                                                                                                                  ifelse(grepl("TurboGrafx", IGN_data$platform, ignore.case = TRUE), "TurboGrafx",
                                                                                                                          ifelse(grepl("WonderSwan", IGN_data$platform, ignore.case = TRUE), "WonderSwan", "Other")))))))))))))

IGN_data %>% group_by(platform, platform_group) %>% 
  summarise(n_distinct(platform))

```


Similar to platform, genre has a multitude of factors which makes reasonable analysis a bit difficult. There are 113 unique genres within the field. I chose to make genre vectors based on an overall description (e.g. Sports, Cards, Action, etc.) to have a cleaner code visual given the quantity of distinct factors. Using those vectors, I created a new variable named genre_group which brought the number of unique genres from 113 to 21. 

```{r group_genre_list}

# 113 different genres need to be grouped into a usable amount of variables 

IGN_data %>% distinct(genre) %>% 
  arrange(genre)

Action <- c("Action", "Action, Adventure", "Action, Compilation", "Action, Editor", "Action, Platformer", "Action, Puzzle", "Action, RPG", "Action, Simulation", "Action, Strategy")

Adventure <- c("Adventure", "Adventure, Adult", "Adventure, Adventure", "Adventure, Compilation", "Adventure, Episodic", "Adventure, Platformer", "Adventure, RPG")

Sports <- c("Baseball", "Sports", "Sports, Action", "Sports, Baseball", "Sports, Compilation", "Sports, Editor", "Sports, Fighting", "Sports, Golf", "Sports, Other", "Sports, Party", "Sports, Racing", "Sports, Simulation")

Cards <- c("Card", "Card, Battle", "Card, Compilation", "Card, RPG")

Compilation <- c("Compilation", "Compilation, Compilation", "Compilation, RPG")

Education <- c("Educational", "Educational, Action", "Educational, Adventure", "Educational, Card", "Educational, Productivity", "Educational, Puzzle", "Educational, Simulation", "Educational, Trivia")

Fighting <- c("Fighting", "Fighting, Action", "Fighting, Adventure", "Fighting, Compilation", "Fighting, RPG", "Fighting, Simulation")
Flight <- c("Flight", "Flight, Action", "Flight, Racing", "Flight, Simulation")

Hunting <- c("Hunting", "Hunting, Action", "Hunting, Simulation")

Music <- c("Music", "Music, Action", "Music, Adventure", "Music, Compilation", "Music, Editor", "Music, RPG")

Pinball <- c("Pinball", "Pinball, Compilation")

Platformer <- c("Platformer", "Platformer, Action", "Platformer, Adventure")

Productivity <- c("Productivity, Action", "Productivity")

Puzzle <- c("Puzzle", "Puzzle, Action", "Puzzle, Adventure", "Puzzle, Compilation", "Puzzle, Platformer", "Puzzle, RPG", "Puzzle, Word Game")

Racing <- c("Racing", "Racing, Action", "Racing, Compilation", "Racing, Editor", "Racing, Shooter", "Racing, Simulation")

RPG <- c("RPG", "RPG, Action", "RPG, Compilation", "RPG, Editor", "RPG, Simulation")

Shooter <- c("Shooter", "Shooter, Adventure", "Shooter, First-Person", "Shooter, Platformer", "Shooter, RPG")

Simulation <- c("Simulation", "Simulation, Adventure")

Strategy <- c("Strategy", "Strategy, Compilation", "Strategy, RPG", "Strategy, Simulation")

Wrestling <- c("Wrestling", "Wrestling, Simulation")

IGN_data$genre_group <- ifelse(grepl(paste(Action, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "Action",
                               ifelse(grepl(paste(Adventure, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "Adventure",
                                      ifelse(grepl(paste(Sports, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "Sports",
                                             ifelse(grepl(paste(Cards, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "Cards",
                                                    ifelse(grepl(paste(Compilation, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "Compilation",
                                                           ifelse(grepl(paste(Education, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "Education",
                                                                  ifelse(grepl(paste(Fighting, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "Fighting",
                                                                         ifelse(grepl(paste(Flight, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "Flight",
                                                                                ifelse(grepl(paste(Hunting, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "Hunting",
                                                                                       ifelse(grepl(paste(Music, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "Music",
                                                                                              ifelse(grepl(paste(Pinball, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "Pinball",
                                                                                                     ifelse(grepl(paste(Platformer, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "Platformer",
                                                                                                            ifelse(grepl(paste(Productivity, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "Productivity",
                                                                                                                   ifelse(grepl(paste(Puzzle, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "Puzzle",
                                                                                                                          ifelse(grepl(paste(Racing, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "Racing",
                                                                                                                                 ifelse(grepl(paste(RPG, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "RPG",
                                                                                                                                        ifelse(grepl(paste(Shooter, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "Shooter",
                                                                                                                                               ifelse(grepl(paste(Simulation, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "Simulation",
                                                                                                                                                      ifelse(grepl(paste(Strategy, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "Strategy",
                                                                                                                                                             ifelse(grepl(paste(Wrestling, collapse = "|"), IGN_data$genre, ignore.case = FALSE), "Wrestling", "Other"))))))))))))))))))))
                                                  
unique(IGN_data$genre_group)
```

After cleaning up the variables that I will be using in my analysis, I wrote the wrangled data to a new file called "ign_clean.csv" for further use later in the course.

```{r cleaned_code_CSV}

write.csv(IGN_data, "ign_clean.csv")

```

















